\documentclass[12pt, a4paper]{article}

% ============================================================
% PACKAGES (verified available on this TeX Live 2025 basic)
% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc}
\usepackage{array}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{authblk}
% ============================================================
% FORMATTING
% ============================================================
\onehalfspacing
\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    citecolor=blue!60!black,
    urlcolor=blue!60!black
}

% Custom commands
\newcommand{\USD}{\text{USD}}
\newcommand{\IDR}{\text{IDR}}
\newcommand{\Mt}{\text{Mt}}
\newcommand{\tCO}{\text{tCO}_2}

% Workaround for missing appendix package
\newcommand{\appendixsection}[1]{\section{#1}}

% ============================================================
% DOCUMENT
% ============================================================
\begin{document}

\title{Stranded Coal and Bank Fragility: Evidence from Indonesia}

\author[1]{Teguh Yudo Wicaksono\thanks{Corresponding author: teguh.wicaksono@uiii.ac.id}}
\author[1]{Rima P Artha}
\affil[1]{Universitas Islam Internasional Indonesia}



\date{\today}

\maketitle

\begin{abstract}
Indonesia is one of the world's largest thermal coal exporters, yet no integrated assessment exists linking coal stranded asset risk to banking sector vulnerability and macroeconomic outcomes. Using firm-level data from 34 listed coal companies and 38 listed banks, we develop a three-model framework to quantify this interrelationship. First, we construct a supply cost curve and estimate scenario-contingent net present values under business-as-usual, 2\textdegree C, and 1.5\textdegree C climate pathways. Second, we build a bipartite bank-coal exposure network from granular loan data and conduct a climate stress test, estimating credit losses and capital adequacy impacts for each bank. Third, we aggregate these results through fiscal, credit, and market channels to estimate the macro-financial impact on GDP. Under a 2\textdegree C scenario, we find aggregate stranded values of \$60.1 billion (90\% CI: \$50.5--72.3 billion), rising to \$99.2 billion (90\% CI: \$83.5--119.0 billion) under a 1.5\textdegree C pathway. These losses generate total bank credit losses of \$1.8 billion (2\textdegree C) and \$3.2 billion (1.5\textdegree C), with the most exposed systemically important bank absorbing losses equivalent to 43\% of annual earnings. The macro-financial impact reaches 0.37\% of GDP under 2\textdegree C and 0.62\% under 1.5\textdegree C, with bank deleveraging as the single largest channel. While no bank breaches regulatory capital thresholds, the concentrated network structure creates idiosyncratic vulnerabilities that require urgent supervisory attention. Our findings highlight the urgency of coordinated climate-financial regulation in commodity-dependent emerging economies.
\end{abstract}

\medskip
\noindent\textbf{Keywords:} Stranded assets; Coal; Banking sector; Climate transition risk; Indonesia; Financial stability

\medskip
\noindent\textbf{JEL Codes:} Q54, G21, G28, E44

\newpage

% ============================================================
% MAIN TEXT
% ============================================================

\input{sections/01_introduction}
\input{sections/02_literature}
\input{sections/03_background}
\input{sections/04_data}
\input{sections/05_model1_stranded_assets}
\input{sections/06_model2_bank_stress}
\input{sections/07_model3_macro_transmission}
\input{sections/08_discussion}
\input{sections/09_conclusion}

\section*{Conflict of interest Declaration}
\noindent
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Funding}
\noindent
This work was supported by \textbf{ViriyaENB} under the grant of \textit{KOMITMEN} at Universitas Islam Internasional Indonesia.

% ============================================================
% REFERENCES
% ============================================================
\bibliographystyle{apalike}
\bibliography{references}

% ============================================================
% APPENDICES
% ============================================================
\appendix
\input{sections/appendix_a_data}
\input{sections/appendix_b_methodology}
\input{sections/appendix_c_results}

% Online supplementary material (for journal submission as separate document)
% \input{sections/supplementary_material}

\end{document}
