%% 02_literature.tex
%% Section 2: Literature Review

\section{Literature Review}
\label{sec:literature}

This paper draws on and contributes to three intersecting literatures: the economics of stranded assets, climate risk and financial stability, and the macro-financial transmission of transition risk. We review each in turn before identifying the specific gap that our integrated framework addresses.

\subsection{Stranded Assets on Carbon}
\label{subsec:stranded_assets}

The intellectual foundations of the stranded assets concept can be traced to the carbon budget literature, which establishes an upper bound on cumulative greenhouse gas emissions compatible with specific temperature targets. \citet{McGlade2015} translate this carbon budget constraint into geographical and fuel-specific terms, estimating that 82\% of global coal reserves, 49\% of gas reserves, and 33\% of oil reserves must remain unextracted to maintain a 50\% probability of limiting warming to 2\textdegree C. \citet{Welsby2021} update this analysis for the more ambitious 1.5\textdegree C target, finding that 89\% of coal reserves globally are ``unextractable,'' with particularly severe implications for coal-exporting nations in Southeast Asia and Africa.

The translation from physical unextractability to economic value destruction is the domain of stranded asset analysis proper. \citet{CarbonTracker2022} popularised the term through their ``carbon bubble'' thesis, arguing that fossil fuel companies are systematically overvalued because markets fail to price the risk that reserves will become economically unviable. \citet{Caldecott2021} provide a comprehensive taxonomy, distinguishing between demand-side stranding (declining coal consumption due to renewable energy competition and climate policy), supply-side stranding (regulatory restrictions on extraction), and cost-side stranding (rising extraction costs as marginal deposits are developed). \citet{VanDerPloeg2020} formalise these dynamics in a resource economics framework, showing that the announcement of future climate policy can trigger immediate asset revaluation as rational agents update expectations.

Empirical quantification of stranded values has proceeded along two tracks. The first employs supply cost curve analysis, ranking fossil fuel deposits by extraction cost and determining which become uneconomic as demand declines or carbon prices rise. \citet{Mendelevitch2019} applies this approach to global steam coal, demonstrating that Indonesian coal is particularly vulnerable owing to its high transport costs to distant markets and competition from lower-cost Australian producers. The second track uses discounted cash flow (DCF) models to value individual firms or projects under alternative climate scenarios. \citet{Edwards2022} apply this methodology to coal-fired power plants, estimating committed emissions and the capital at risk from early retirement. \citet{Pfeiffer2018} extend the analysis to planned infrastructure, finding that committed emissions from existing and planned power plants alone could exhaust the 1.5\textdegree C carbon budget.

Our work is closely related to \citet{sonny2016}, who provide the first Indonesia-specific quantification of coal stranded assets. Using a cash cost approach applied to 16 listed coal companies representing approximately 60\% of national production, they construct a supply cost curve for the Indonesian coal industry and estimate aggregate stranded values ranging from \$58 billion to \$99 billion, depending on the discount rate (6--8\%), assumed mine life (15--20 years), and the magnitude of coal price decline (30--40\%). At their central specification (8\% discount rate, 20-year life, 40\% price decline from the 2014 benchmark), the aggregate stranded value reaches \$86 billion. Their firm-level results reveal substantial heterogeneity: low-cost producers such as Kideco face smaller per-unit losses, while higher-cost firms such as ITMG and Bayan Resources are pushed to the margin of viability. Importantly, \citet{sonny2016} argue that Indonesia may temporarily delay asset stranding by aggressively expanding domestic coal consumption through the Domestic Market Obligation (DMO) and subsidised mine-mouth power plants. They also note that the fiscal impact on state revenues might be less severe than anticipated due to historically poor collection of coal rents.

While \citet{sonny2016} provide valuable firm-level estimates and highlight the fiscal dimension, their analysis does not extend to the banking sector or to macroeconomic transmission channels. Their cash cost methodology, while intuitive, does not incorporate scenario-contingent DCF modelling with Monte Carlo uncertainty quantification, which limits the ability to produce probabilistic stranded value estimates under alternative climate pathways.

All in all, a general limitation of this literature is its predominantly physical or sectoral focus. Most studies quantify the volume of unburnable reserves or the value at risk within the fossil fuel sector, but do not trace how these losses propagate to the financial system or the broader economy. This paper addresses that limitation by linking firm-level stranding estimates to bank exposures and macroeconomic outcomes.

\subsection{Climate Risk and Financial Stability}
\label{subsec:climate_risk}

The recognition that climate change poses risks to financial stability has accelerated rapidly since Mark Carney's influential ``tragedy of the horizon'' speech in 2015 and the establishment of the Network for Greening the Financial System (NGFS) in 2017. \citet{Battiston2017} provide one of the first quantitative assessments, stress-testing the European financial system against a sudden fossil fuel price decline and finding that direct exposures to carbon-intensive sectors represent a material but manageable risk for most European banks. However, they note that indirect exposures through interbank lending and derivatives could amplify losses substantially.

Subsequent climate stress tests have been conducted by the Dutch central bank \citep{Vermeulen2018}, the Bank of England \citep{BoE2019}, and the European Central Bank \citep{ECB2022}. These exercises reveal several common findings: transition risk is concentrated in specific sectors (energy, transport, heavy industry); bank losses depend critically on whether the transition is orderly or disorderly; and second-round effects, including fire sales, counterparty defaults, and funding market disruptions, can substantially amplify first-round losses. The IMF has extended this framework to emerging markets, including a Financial Sector Assessment Program for Indonesia \citep{IMF2024Indonesia} that identifies coal exposure as a material risk but stops short of a full climate stress test.

On the asset pricing side, \citet{BoltonKacperczyk2021} document a carbon risk premium in equity markets, showing that firms with higher carbon emissions earn higher stock returns, consistent with investors demanding compensation for transition risk. \citet{Dietz2016} estimate the climate value-at-risk of global financial assets at \$2.5 trillion under a business-as-usual scenario, with a fat-tailed distribution that implies non-trivial probability of losses exceeding \$24 trillion. \citet{Daumas2024} provides a comprehensive survey of this literature, noting that while theoretical understanding of climate-financial linkages has advanced considerably, empirical evidence grounded in actual firm-level data remains scarce, particularly for emerging markets.

Our contribution to this literature is twofold. First, we construct a bipartite bank-coal exposure network from actual loan data rather than relying on sectoral aggregates, allowing us to identify specific bank-firm channels of risk transmission. Second, we conduct what is, to our knowledge, the first bottom-up climate stress test of the Indonesian banking system that traces losses from individual coal company stranding through identified credit exposures to bank-level capital adequacy impacts.

\subsection{Macro-Financial Transmission of Transition Risk}
\label{subsec:macro_transmission}

The third strand of literature examines how climate transition risk propagates from the financial sector to the macroeconomy. The theoretical foundations draw on the financial accelerator framework of \citet{Bernanke1999}, which demonstrates how shocks to firm net worth can be amplified through credit market frictions, generating business cycle fluctuations that exceed the magnitude of the initial disturbance. \citet{Carattini2023} extend this framework to climate policy, showing that transition risk can trigger a ``green paradox in reverse,'' where the announcement of future carbon pricing causes immediate asset revaluations that tighten financial conditions and depress output.

\citet{Mercure2018} provide the most comprehensive quantification of the macroeconomic impact of stranded assets, using the E3ME macro-econometric model to estimate GDP losses of 1--4\% for fossil fuel exporting nations under a 2\textdegree C scenario. Their analysis identifies two key transmission channels: the direct loss of resource rents and the indirect effects through financial markets and investment. \citet{Grippa2019} formalise a three-channel framework (fiscal, credit, and market) that has since been adopted by the IMF for emerging market assessments, though empirical calibration remains limited.

The fiscal dimension is particularly salient for resource-dependent economies. \citet{Arellano2008} and \citet{GennaiGoldin2014} demonstrate how commodity price declines can trigger sovereign stress through falling tax revenues and rising debt-to-GDP ratios, with feedback loops to the banking sector through sovereign-bank nexus channels \citep{Acharya2017}. For Indonesia, coal royalties and corporate taxes from the mining sector contributed approximately 4.5\% of central government revenue in 2022, while state-owned enterprises such as PT Bukit Asam (PTBA) provide additional dividend income.

The credit channel operates through the bank lending mechanism identified by \citet{PeekRosengren2000} and \citet{GambacortaMarques2011}: when banks suffer credit losses, they reduce lending to the broader economy, with multiplier effects that depend on the degree of financial intermediation and the availability of alternative funding sources. In Indonesia's bank-dominated financial system, where bank credit finances approximately 34\% of GDP, this channel is likely to be more potent than in market-based financial systems.

Our paper contributes to this literature by providing the first empirically calibrated, bottom-up estimate of the macro-financial impact of coal stranding in an emerging economy. Rather than relying on top-down macro models or stylised assumptions about financial linkages, we build from firm-level stranding data through identified bank exposures to aggregate macroeconomic transmission, allowing us to decompose the total impact across fiscal, credit, and market channels.

\subsection{Indonesia: Coal, Banking, and the Transition Gap}
\label{subsec:indonesia_gap}

Indonesia-specific research on climate transition risk has grown in recent years but remains fragmented. \citet{IESR2023} provides the most comprehensive assessment of coal as stranded assets in Indonesia, using a combination of reserve analysis, scenario modelling, and qualitative banking assessment. Their report identifies 13 of 34 coal companies as highly vulnerable to stranding under a 1.5\textdegree C scenario and notes that four major banks hold significant coal exposure, but does not quantify credit losses or macroeconomic impacts. The IMF's Financial Sector Assessment Program for Indonesia \citep{IMF2024Indonesia, IMF2025IndonesiaStress} incorporates climate risk as a supplementary consideration but focuses primarily on physical risks (flooding, sea-level rise) rather than transition risk from coal stranding.

\citet{Rishanty2023} provide the most recent and methodologically ambitious Indonesia-specific assessment, combining mine-level quantitative modelling (developed in collaboration with Willis Towers Watson) with qualitative stakeholder analysis. Using granular mine-level data covering total Indonesian coal production, they estimate the external climate transition value-at-risk to the coal sector at \$206 billion (NPV, 2023) under a Well Below 2\textdegree C (WB2C) scenario over 2023--2050. This risk is distributed across investors in coal firms (\$35 billion), the government (\$51 billion), and the coal supply chain (\$120 billion). Their analysis highlights that approximately 85\% of Indonesia's coal exports are destined for countries with net-zero commitments, creating substantial external demand risk. They also estimate that 51,000 fewer mining jobs would exist by 2030 under WB2C compared to BAU, with mines in East and South Kalimantan most severely affected. Critically, \citet{Rishanty2023} emphasise the role of banks and the central bank in financing local economic transitions, but do not construct a bank-level stress test or quantify credit losses through identified exposure channels.

\citet{Burke2023} examine Indonesia's energy policy landscape, documenting the deep structural dependence on coal for both electricity generation and export revenues. \citet{Makhijani2015} quantify fossil fuel subsidies in Indonesia, which further entrench coal dependence by artificially lowering the cost of coal-fired electricity relative to renewables. \citet{JakobSteckel2016} situate Indonesia within the broader political economy of energy transitions in developing countries, noting that concentrated coal interests and dispersed transition benefits create formidable barriers to policy change.

To summarise, while prior studies have quantified the magnitude of Indonesian coal stranded assets at the firm level \citep{sonny2016} or the mine level \citep{Rishanty2023}, and have qualitatively noted the banking sector implications \citep{IESR2023}, no existing study provides an integrated, quantitative assessment of how coal stranding transmits through the banking sector to the macroeconomy. Table~\ref{tab:comparison} in Section~\ref{sec:discussion} provides a detailed comparison of our approach with prior work, highlighting the novelty of our firm-level, bottom-up, three-model framework.
