%% 09_conclusion.tex
%% Section 9: Conclusion

\section{Conclusion}
\label{sec:conclusion}

This paper develops the first integrated framework for assessing how coal stranded asset risk propagates through the banking sector in Indonesia, the world's largest thermal coal exporter. Using firm-level data from 35 listed coal companies and 38 listed banks, we construct a two-model framework that links bottom-up stranded asset valuations to bank-level credit losses through a granular bipartite exposure network.

Our key findings are as follows. First, under a 2\textdegree C-aligned climate transition, aggregate coal stranded values reach \$79.7 billion (90\% CI: \$72.7--88.0 billion from Monte Carlo simulation with 10,000 draws), rising to \$128.3 billion (90\% CI: \$117.1--141.0 billion) under a 1.5\textdegree C pathway. Stranding risk is highly concentrated: the top five companies account for over 75\% of total stranded value, with important implications for targeted just-transition support. Second, the banking sector stress test reveals total credit losses of \$1.8 billion under the 2\textdegree C scenario and \$3.1 billion under 1.5\textdegree C, with losses concentrated in a small number of large banks. While no bank breaches regulatory capital thresholds, the most exposed D-SIB (BMRI) absorbs losses equivalent to 42\% of annual earnings, highlighting the idiosyncratic vulnerability created by concentrated coal lending. The bipartite network structure further reveals correlated default risk that limits intra-sector diversification, amplifying systemic exposure beyond what individual exposure limits can capture.

These findings contribute to the literature in two ways. To the stranded assets literature, we contribute the most granular firm-level valuation of coal stranding in any single country, with Monte Carlo uncertainty quantification that moves beyond deterministic point estimates. To the climate stress testing literature, we contribute the first bottom-up stress test of the Indonesian banking system using actual bank-coal credit relationships rather than sectoral aggregates, demonstrating that bipartite network analysis can identify idiosyncratic vulnerabilities invisible in aggregate data.

The policy implications are clear. Indonesia's regulators, OJK and Bank Indonesia, should integrate coal transition risk into supervisory stress testing, enhance fossil fuel exposure disclosure requirements, and consider climate-adjusted capital requirements for coal-exposed banks. The JETP's success depends on managing the financial vulnerabilities we document: an orderly, well-signalled transition that allows banks and firms to adjust gradually will minimise the credit channel amplification that our analysis identifies as the dominant transmission mechanism. Conversely, a disorderly transition risks triggering the bank deleveraging spiral that converts firm-level stranding into a macroeconomic headwind.

More broadly, the coal-banking-sovereign nexus we document is likely replicated in other coal-dependent emerging economies, including South Africa, India, Colombia, and the Philippines. Our framework provides a replicable template for country-specific assessments, and we encourage future research to extend the analysis to these contexts. Additional directions for future work include: (1) incorporating dynamic transition pathways using agent-based or DSGE models that can capture the interaction between transition speed, capital expenditure decisions, and loan maturity profiles; (2) modelling second-round effects through interbank contagion and sovereign-bank feedback loops, which our analysis identifies as likely material but does not quantify; (3) extending the analysis to the full coal value chain, including coal-fired power utilities, logistics providers, and port operators, whose bank exposures represent transition risk beyond the mining sector; (4) incorporating sovereign stress into the framework, to assess the feedback loop between fiscal revenue losses, sovereign credit risk, and bank balance sheets in an economy where state-owned banks are major creditors of the sovereign; and (5) conducting a comparative cross-country study of the coal-banking-sovereign nexus across South Africa, India, Colombia, and the Philippines, which would allow systematic identification of the structural factors that determine whether coal transition risk is primarily a banking problem, a fiscal problem, or a macroeconomic stability problem. Each of these extensions builds directly on the data infrastructure and methodological framework we have established here.

The urgency of this research agenda is underscored by an important asymmetry in the timing of transition risk. The financial losses from coal stranding are already accumulating in bank balance sheets, as coal prices normalise from their 2022--2023 supercycle peaks and investors reassess the long-run demand outlook. Yet most supervisory frameworks, including OJK's, do not yet require forward-looking climate scenario analysis in internal credit risk assessments, meaning that current CKPN levels likely understate the transition-adjusted expected loss on coal portfolios. Acting before coal prices decline further --- rather than after the sector reaches financial distress --- allows for orderly provisioning, negotiated transition arrangements, and the preservation of banking system resilience. The window for this proactive approach is narrowing with each year that climate policy strengthens and renewable costs fall. Our analysis demonstrates that the tools exist to make this assessment rigorously and that the financial and institutional stakes are high enough to make doing so a priority for Indonesia's financial regulators and their counterparts across coal-dependent emerging economies.
